Honors Draft

### 2020-02-10 chapters/article-draft.tex
checking to see if pdf will compile

---
title: "Honors Draft"
author: "Raven McKnight"
date: "2/9/2020"
output:
  pdf_document:
    number_sections: true
  html_document:
    theme: paper
    toc: yes
    toc_float: yes
abstract: "abstract"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

Public transit ridership has been in decline around the country for years, but particularly since the mid 2010s (). Following a spike in ridership related to the 2008 housing market crash, ridership has been steadily trending downwards in most metropolitan areas (). Bus ridership is particularly affected compared to more "desirable" light rail and commuter rail. Transit agencies are naturally interested in understanding the specific causes of these declines, as well as more generally being able to predict transit demand throughout space and time.

@@

It is somewhat rare to see more traditional statistical models applied to transit ridership. In the context of predicting future ridership, the above methods are often sufficient. Additionally, they are the industry standard and allow for the integration of expert planning knowledge. However, in the context of understanding *past* ridership, we can apply more explanatory methods. Specifically, we can use hierarchical spatial Bayesian models to incorporate many covariates in addition to the spatial structure that necessarily underpins transit ridership (ie, boardings can only occur at existing bus stops).

Hierarchical spatial models allow us to understand the patterns of ridership within a city with great granularity. Some existing literature explores the decline of transit ridership *between* cities (). Few studies have been conducted on the various patterns of ridership *within* a single metropolitan area. Additionally, few studies incorporate many demographic predictors. Naturally, we cannot collect demographic information for every rider of a transit system. Instead, we can use demographic predictors associated with spatial units such as census block groups to explore the question of *who* rides -- or, at the very least, *where* do they ride.

# Background

## Metro Transit

@@

## Transit Market Areas

@@

$$
\begin{aligned}
\text{Transit Market Index } = & \text{ 0.64 * Population Density + 0.20 * Employment Density +} \\
& \text{0.23 * Intersection Density + 0.11 * Automobile Availability}
\end{aligned}
$$

where each predictor is logged and scaled by developed land acreage per census block group. Here, automobile availability refers to the number of adults over age 16 less the total number of automobiles available in a block group (scaled by developed land acreage). There is an additional indicator variable, omitted in this notation, for the census block group containing the MSP International Airport. For more documentation of the official Transit Market Areas, refer to the TPPs [Appendix G](https://metrocouncil.org/Transportation/Planning-2/Key-Transportation-Planning-Documents/Transportation-Policy-Plan/The-Adopted-2040-TPP-(1)/Final-2040-Transportation-Policy-Plan/2040-TPP-Appendix-G-Transit-Design-and-Perf-Standa.aspx).

@@

The Transit Market Index values are geographically smoothed to create more-or-less concentric Transit Market Areas. The linear regression is an intuitive way to think about transit ridership across space. The four predictors used are common-sense indicators of transit ridership: high population and employment density are characteristic of trip origins and destinations, automobile availability is reasonably assumed to be related to transit ridership, and intersection density is a proxy for the "walkability" of an area. However, the exsiting Transit Market Area model may be over-simplifying the complex question of transit ridership.

# Data

## Metro Transit Data

@@

**Automatic Passenger Count (APC)** data is reported every time a bus opens its doors. Two beams of light detect movement through the doors of the bus and counts the number of passengers getting on and off at each bus stop. Naturally, this data source is flawed: sometimes, vehicles fail to report data, and the sensors can be easily tricked. Someone getting off the bus pulling a suitcase behind them, for example, would likely be counted as two passengers exiting. However, APC is the most granular ridership data available, giving us the most control over our spatial aggregations of ridership. This is the primary data source used in this analysis.

**Automatic Vehicle Location (AVL)** data is similarly reported by in-service vehicles (approximately) every 8 seconds. The vehicle reports in GPS location continuously while in service. We use AVL data to adjust for missing APC data.

**Metro Transit's Schedule** as *planned*, rather than as *run*, is also used to correct missing APC data.

This analysis began by pulling all APC, AVL, and schedule data from 2015-2018. The initial data pull consisted of approximately 496 *billion* rows of data. The analysis presented here focuses on 2017 data. Figure *___* shows the average weekday boardings by census block group in 2017.

@@

### Data Interpolation

In theory, we have APC data for each run bus trip in 2017. However, we know that, for a variety of reasons, this is untrue. The automatic data reporting technology on Metro Transit buses is flawed and can fail to report data unpredictably. Additionally, the trips missing data may not be random: all buses of one "series", for example, may fail to report at once. This could lead to the APC dataset understimating boardings on a particular set of routes or locations.T herefore, we use the three data sets described above to create a more complete augmented APC data set.

@@

1. Compares scheduled trips to trips we have APC data for. If a trip was scheduled but has no APC records, we consider that trip missing. 
2. For missing trips, check AVL data. If there are no AVL records, we consider the trip cut (ie, we assume the trip was not run). 
3. For missing trips with AVL data, we estimate boardings for that trip by taking the average number of boardings for trips of that nature (trips on the same route, day of week, season, location, etc).

### Data Aggregation

@@

## Other data sources

@@

Note that `beta[3]` is a measure of walkability. A 10-minute walk isochrone, generated using OpenTripPlanner, shows the area a person on foot can cover in 10 minutes on the existing street network. In theory, we expect more walkable areas, such as the downtown Saint Paul and Minneapolis, to have larger walk-isochrones. Unfortunately, the metric can be fooled by rural areas where you can walk uninterruped down a single roadway, for example. An improved "walkability" metric could improve models of transit ridership.

@@

# Models

## Stan

@@

* $\hat{R}$ or `Rhat` indicates how well the Markov chains have mixed by comparing between- and within-chain estimates. At convergence, $\hat{R} = 1$. When $\hat{R} > 1$, the Markov chains have not mixed well and the model likely needs to be run for more iterations.  
* `n_eff` is the estimated Effective Sample Size (ESS). ESS is the number of independent samples our sample is equivalent to. In other words, if `n_eff` = 1,000, we expect to gain the same amount of information from our sample as we would from 1,000 independent samples. In a "perfect" scenario, `n_eff` should equal the actual number of samples. When `n_eff` is significantly below the actual number of samples, there may be a problem with the model and posterior estimates may be biased. 
* **Divergent transitions** occur when the simulated Hamiltonian departs from the actual Hamiltonian trajectory. In general, if stan gives any divergent transition warnings, we should consider the posterior estimates unreliable.  
* `max_treedepth` is a parameter we set when fitting a model in Stan. Treedepth refers to the number of steps the sampler has to take at each iteration. The default is 10. For extremely "curved" or complex posteriors, the treedepth may be much higher. When `max_treedepth` is exceeded, the sampler quits prematurely. Therefore, if we have *many* iterations exceeding `max_treedepth`, posterior estimates may be unreliable. Additionally, setting high treedepth causes the sampler to perform much more slowly. 
* `adapt_delta` is also a sampling parameter; it corresponds to the Mean Metropolis Acceptance Rate (see ____). The default is 0.8. Setting a higher `adapt_delta` slows computation time but *can* eliminate "false alarm" divergent transitions.

## Poisson Regression

For each block group $i = 1, 2, 3, ..., 1495$ with any ridership, the "baseline" model is a simple Poisson regression. Poisson regressions are a type of generalized linear model with the natural log as its link function. For ridership $Y_i$ and covariates $x_1, ..., x_k$ discussed above, the Poisson regression is written

$$
\begin{aligned}
Y_i & \sim \text{Poisson}(E_i\lambda_i) \\
\eta_i  = log(\lambda_i) & = \beta_0 + \sum_{k=1}^{K}x_k^T\beta_k \\
\beta_0, \beta &\sim \text{Normal}(0, 1)
\end{aligned}
$$

where $\beta_0$ is the intercept and $E_i$ is an offset term.

The offset $E_i$ is often termed "exposure" in applications such as disease risk modeling where the offset may be the number of observed cases in a previous year, for example. More technically, the offset term scales the Poisson output to be a rate rather than a count. This is appropriate when observations $i$ have different potentials for response $Y$. For example, a county with higher population will naturally have a higher count of patients with asthma than a county with a smaller population. Mathematically, the offset is a covariate with parameter set equal to 1. Recall, the Poisson regression assumes we can model the mean of response $Y$, $\bar{Y}$ with a combination of linear predictors:

$$
log(\bar{Y}) = \beta_0 + \sum_{k=1}^{K}x_k^T\beta_k \\
$$
Therefore, if we wish to model the rate $Y/E$, the equation is rewritten

$$
\begin{aligned}
log(\bar{Y}/E) & = \beta^{'}_0 + \sum_{k=1}^{K}x_k^T\beta^{'}_k \\
log(\bar{Y}) & = log(E) + \beta^{'}_0 + \sum_{k=1}^{K}x_k^T\beta^{'}_k
\end{aligned}
$$

In this study, we define $E_i$ to equal the number of times a bus stops in block group $i$. This allows us to control for block groups with more or less supply.

In the basic Poisson regression, we often give $\beta_0$ and $\beta$ vague priors such as Normal($0, 1$). The exposure term $E_i$ comes from the data and does not recieve a prior.

### Model 1

We will call this baseline Poisson regression Model 1. Model 1 uses all 18 covariates with simple Normal(0, 1) priors and no overdispersion parameters. The Stan program used to fit Model 1, as well as the other four models, can be found in the appendix. Model 1 fits quickly and efficiently. The Stan output for $\beta_0$ and $\beta$ is below.

@@

Model 1 produces coefficient estimates with tight confidence intervals; few of the $\beta_k$ 95% intervals cross zero. Based on initial diagnostics and precise parameter estimates, Model 1 appears to be a "good" fit. In this way, Model 1 illustrates the importance of diagnostics beyond the default Stan output.

@@

## Overdispersed Poisson Regression

The most obvious assumption made in a Poisson regression is that response variable $Y$ follows a Poisson distribution. In addition to requiring integer counts, this assumption requires that $\text{Var}(Y) = \text{E}(Y)$. This is often not true in practice. In the case of the Metro Transit ridership data, $\text{Var}(Y) >>\text{E}(Y)$. In this case, we call the unexpectedly large variance "overdispersion" or "extra-Poisson variance."

@@

$$
\begin{aligned}
\eta_i = log(\lambda_i) & = \beta_0 + \sum_{k=1}^{K}x_k^T\beta_k + \theta\sigma \\
\theta &\sim \text{Normal}(0, 1) \\
\sigma &\sim \text{Normal}(0, 5)
\end{aligned}
$$

where $\theta$ is a set of heterogenous random effects. In practice, $\theta$ is scaled by its variance parameter $\sigma$ to allow eassier fitting in Stan. The addition of this set of random effects improves fits on data with overdispersion. With sufficiently vague priors, $\theta$ and $\sigma$ can account for most or all of the overdispersion not accounted for by $\beta_0 + \sum_{k=1}^{K}x_k^T\beta_k$.

### Model 2

Model 2 builds upon Model 1 by incorporating the set of random effects described above. Model 2 is *significantly* more computationally expensive to fit than Model 1. This is due to relatively large autocorrelation between each iteration, visible in the traceplots below (Figure ___). This autocorrelation made it necessary to run Model 2 on four chains for 20,000 iterations each. Despite this, `n_eff` is significantly lower for Model 2 than Model 1. Model 2 also experienced 2 transitions exceeding maixmum treedepth: not a problem in terms of model fit, but a signal that the posterior is more challenging to fit. Reparameterization could likely reduce autocorrelation between iterations.

The parameter estimates for $\beta_0$ and $\beta$ for Model 2 are below. Notably, the direction and magnitude of the coefficient estimates are similar to those produced by Model 1. However, the coverage intervals are much wider. Based on "test runs" of Model 2, we expect increasing the number of iterations to decrease the width of these coverage intervals (we observe that the converage interval is proportional to the number of iterations in this setting).

@@

<!-- \begin{figure}[H] -->
<!--   \centering -->
<!--   \includegraphics[width = 6in]{ppcheckmod2.png} -->
<!--   \caption{Posterior predictive check for Model 1.} -->
<!-- \end{figure} -->

Because we gave $\theta$ and $\sigma$ vague priors, we can see a *much* better fit in the posterior predictive checks for Model 2. One *could* stop analysis here. However, the addition of $\theta \sigma$ only improves the *fit* of the model, not our understanding of the data. Therefore, for this study, we will decompose $\theta \sigma$ into spatial and non-spatial random effects. First, however, we will implement a Bayesian variable selection process to determine which, if any, of our covariates $x_k$ should be removed from the analysis.

## Horseshoe Priors and Variable Selection

In the baseline model, we give coefficients $\beta_1, ..., \beta_k$ simple normal priors. With a large set of possible covariates, however, we may reasonably expect some of the $\beta_k$ coefficients to be equal to zero -- in other words, that some of the parameters $x_k$ have no effect on response $Y$. In frequentist applications, we might use a method such as the LASSO (Least Absolute Shrinkage and Selection Operator) to identify relevant variables. There are several Bayesian alternatives for variable selection with generally consist of applying particular priors to $\beta$.

### Spike and Slab - needs graphics

In the Bayesian setting, there are two primary methods for variable selection. The first is called the "spike-and-slab" prior (Mitchell and Beauchamp, 1988; George and McCulloch, 1999) and has often been considered the "gold standard" for sparse Bayesian regression, or variable selection (Piironen and Vehtari, 2017). This prior is often expressed as

$$
\begin{aligned}
\beta_j | \lambda_j, c, \epsilon & \sim \lambda_j \text{Normal}(0, c^2) + (1-\lambda_j)\text{Normal}(0, \epsilon^2) \\
\lambda_j & \sim \text{Bernoulli}(\pi)
\end{aligned}
$$

for $j = 1, 2, ..., J$ where $\lambda_j \in {0, 1}$ indicates whether $\beta_j$ is from the "spike" (ie, near zero) or the "slab" (ie, nonzero). The "spike" is the area where most of the prior density for $\beta$ is centered (generally around 0) while the "slab" refers to the low-density extent of the prior. The spike-and-slab prior encourages all $\beta_j$ towards zero; only $\beta_j$ sufficiently far from 0 will be estimated to be from the slab.

The primary challenge with the spike-and-slab (and other Bayesian shrinkage methods) is the selction of priors. Priors must be set for the width of the slab $c$ and the prior inclusion probability $\pi$. Here, $\pi$ reflects our prior understanding of the sparsity of $\beta$. In practice, we rarely have strong prior knowledge of the number of predictors we expect to be distinguishable from zero. This can make the implementation of sparse regression more challenging than, say, the frequentist LASSO.

### Horseshoe - needs graphics

The second method for Bayesian sparse regression is to give $\beta$ some continuous *sparsity inducing prior*. **Van Erp et al (2019)** provide an introduction to many such continous priors. Ridge regression and the Bayesian LASSO are two of the most approachable priors, but all sparsity inducing priors utilize the same logic as the spike-and-slab: give $\beta$ a prior with *most* of its mass near 0, to shrink irrelevant coefficients, and the rest of its mass "far" from 0. Identifying "far" depends more or less on the specific data and model in question.

The horseshoe prior, proposed by **Carvalho et al, 2010** is a popular choice in Bayesian literature, in part because it is similar to the "gold standard" spike-and-slab. Following the notation of **Vehtari & Piironen**, the horseshoe prior can be expressed

$$
\begin{aligned}
\beta_k \text{ | } \lambda_k, \tau & \sim \text{Normal}(0, \tau^2\lambda_k^2) \\
\lambda_k & \sim \text{Half-Cauchy}(0, 1)
\end{aligned}
$$

for $k = 1, 2, ..., K$. The horseshoe prior is so-named because for fixed values $\tau = \lambda_k = 1$, the prior resemebles a Beta(1/2, 1/2) distribution, or a horseshoe.

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Honors Draft},
            pdfauthor={Raven McKnight},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Honors Draft}
\author{Raven McKnight}
\date{2/9/2020}

\begin{document}
\maketitle
\begin{abstract}
abstract
\end{abstract}

\section{Introduction}\label{introduction}

Public transit ridership has been in decline around the country for
years, but particularly since the mid 2010s (). Following a spike in
ridership related to the 2008 housing market crash, ridership has been
steadily trending downwards in most metropolitan areas (). Bus ridership
is particularly affected compared to more ``desirable'' light rail and
commuter rail. Transit agencies are naturally interested in
understanding the specific causes of these declines, as well as more
generally being able to predict transit demand throughout space and
time.

@@

It is somewhat rare to see more traditional statistical models applied
to transit ridership. In the context of predicting future ridership, the
above methods are often sufficient. Additionally, they are the industry
standard and allow for the integration of expert planning knowledge.
However, in the context of understanding \emph{past} ridership, we can
apply more explanatory methods. Specifically, we can use hierarchical
spatial Bayesian models to incorporate many covariates in addition to
the spatial structure that necessarily underpins transit ridership (ie,
boardings can only occur at existing bus stops).

Hierarchical spatial models allow us to understand the patterns of
ridership within a city with great granularity. Some existing literature
explores the decline of transit ridership \emph{between} cities (). Few
studies have been conducted on the various patterns of ridership
\emph{within} a single metropolitan area. Additionally, few studies
incorporate many demographic predictors. Naturally, we cannot collect
demographic information for every rider of a transit system. Instead, we
can use demographic predictors associated with spatial units such as
census block groups to explore the question of \emph{who} rides -- or,
at the very least, \emph{where} do they ride.

\section{Background}\label{background}

\subsection{Metro Transit}\label{metro-transit}

@@

\subsection{Transit Market Areas}\label{transit-market-areas}

@@

\[
\begin{aligned}
\text{Transit Market Index } = & \text{ 0.64 * Population Density + 0.20 * Employment Density +} \\
& \text{0.23 * Intersection Density + 0.11 * Automobile Availability}
\end{aligned}
\]

where each predictor is logged and scaled by developed land acreage per
census block group. Here, automobile availability refers to the number
of adults over age 16 less the total number of automobiles available in
a block group (scaled by developed land acreage). There is an additional
indicator variable, omitted in this notation, for the census block group
containing the MSP International Airport. For more documentation of the
official Transit Market Areas, refer to the TPPs
\href{https://metrocouncil.org/Transportation/Planning-2/Key-Transportation-Planning-Documents/Transportation-Policy-Plan/The-Adopted-2040-TPP-(1)/Final-2040-Transportation-Policy-Plan/2040-TPP-Appendix-G-Transit-Design-and-Perf-Standa.aspx}{Appendix
G}.

@@

The Transit Market Index values are geographically smoothed to create
more-or-less concentric Transit Market Areas. The linear regression is
an intuitive way to think about transit ridership across space. The four
predictors used are common-sense indicators of transit ridership: high
population and employment density are characteristic of trip origins and
destinations, automobile availability is reasonably assumed to be
related to transit ridership, and intersection density is a proxy for
the ``walkability'' of an area. However, the exsiting Transit Market
Area model may be over-simplifying the complex question of transit
ridership.

\section{Data}\label{data}

\subsection{Metro Transit Data}\label{metro-transit-data}

@@

\textbf{Automatic Passenger Count (APC)} data is reported every time a
bus opens its doors. Two beams of light detect movement through the
doors of the bus and counts the number of passengers getting on and off
at each bus stop. Naturally, this data source is flawed: sometimes,
vehicles fail to report data, and the sensors can be easily tricked.
Someone getting off the bus pulling a suitcase behind them, for example,
would likely be counted as two passengers exiting. However, APC is the
most granular ridership data available, giving us the most control over
our spatial aggregations of ridership. This is the primary data source
used in this analysis.

\textbf{Automatic Vehicle Location (AVL)} data is similarly reported by
in-service vehicles (approximately) every 8 seconds. The vehicle reports
in GPS location continuously while in service. We use AVL data to adjust
for missing APC data.

\textbf{Metro Transit's Schedule} as \emph{planned}, rather than as
\emph{run}, is also used to correct missing APC data.

This analysis began by pulling all APC, AVL, and schedule data from
2015-2018. The initial data pull consisted of approximately 496
\emph{billion} rows of data. The analysis presented here focuses on 2017
data. Figure *\_\_\_* shows the average weekday boardings by census
block group in 2017.

@@

\subsubsection{Data Interpolation}\label{data-interpolation}

In theory, we have APC data for each run bus trip in 2017. However, we
know that, for a variety of reasons, this is untrue. The automatic data
reporting technology on Metro Transit buses is flawed and can fail to
report data unpredictably. Additionally, the trips missing data may not
be random: all buses of one ``series'', for example, may fail to report
at once. This could lead to the APC dataset understimating boardings on
a particular set of routes or locations.T herefore, we use the three
data sets described above to create a more complete augmented APC data
set.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Compares scheduled trips to trips we have APC data for. If a trip was
  scheduled but has no APC records, we consider that trip missing.
\item
  For missing trips, check AVL data. If there are no AVL records, we
  consider the trip cut (ie, we assume the trip was not run).
\item
  For missing trips with AVL data, we estimate boardings for that trip
  by taking the average number of boardings for trips of that nature
  (trips on the same route, day of week, season, location, etc).
\end{enumerate}

\subsubsection{Data Aggregation}\label{data-aggregation}

@@

\subsection{Other data sources}\label{other-data-sources}

@@

Note that \texttt{beta{[}3{]}} is a measure of walkability. A 10-minute
walk isochrone, generated using OpenTripPlanner, shows the area a person
on foot can cover in 10 minutes on the existing street network. In
theory, we expect more walkable areas, such as the downtown Saint Paul
and Minneapolis, to have larger walk-isochrones. Unfortunately, the
metric can be fooled by rural areas where you can walk uninterruped down
a single roadway, for example. An improved ``walkability'' metric could
improve models of transit ridership.

@@

\section{Models}\label{models}

\subsection{Stan}\label{stan}

@@

\begin{itemize}
\tightlist
\item
  \(\hat{R}\) or \texttt{Rhat} indicates how well the Markov chains have
  mixed by comparing between- and within-chain estimates. At
  convergence, \(\hat{R} = 1\). When \(\hat{R} > 1\), the Markov chains
  have not mixed well and the model likely needs to be run for more
  iterations.\\
\item
  \texttt{n\_eff} is the estimated Effective Sample Size (ESS). ESS is
  the number of independent samples our sample is equivalent to. In
  other words, if \texttt{n\_eff} = 1,000, we expect to gain the same
  amount of information from our sample as we would from 1,000
  independent samples. In a ``perfect'' scenario, \texttt{n\_eff} should
  equal the actual number of samples. When \texttt{n\_eff} is
  significantly below the actual number of samples, there may be a
  problem with the model and posterior estimates may be biased.
\item
  \textbf{Divergent transitions} occur when the simulated Hamiltonian
  departs from the actual Hamiltonian trajectory. In general, if stan
  gives any divergent transition warnings, we should consider the
  posterior estimates unreliable.\\
\item
  \texttt{max\_treedepth} is a parameter we set when fitting a model in
  Stan. Treedepth refers to the number of steps the sampler has to take
  at each iteration. The default is 10. For extremely ``curved'' or
  complex posteriors, the treedepth may be much higher. When
  \texttt{max\_treedepth} is exceeded, the sampler quits prematurely.
  Therefore, if we have \emph{many} iterations exceeding
  \texttt{max\_treedepth}, posterior estimates may be unreliable.
  Additionally, setting high treedepth causes the sampler to perform
  much more slowly.
\item
  \texttt{adapt\_delta} is also a sampling parameter; it corresponds to
  the Mean Metropolis Acceptance Rate (see \_\_\_\_). The default is
  0.8. Setting a higher \texttt{adapt\_delta} slows computation time but
  \emph{can} eliminate ``false alarm'' divergent transitions.
\end{itemize}

\subsection{Poisson Regression}\label{poisson-regression}

For each block group \(i = 1, 2, 3, ..., 1495\) with any ridership, the
``baseline'' model is a simple Poisson regression. Poisson regressions
are a type of generalized linear model with the natural log as its link
function. For ridership \(Y_i\) and covariates \(x_1, ..., x_k\)
discussed above, the Poisson regression is written

\[
\begin{aligned}
Y_i & \sim \text{Poisson}(E_i\lambda_i) \\
\eta_i  = log(\lambda_i) & = \beta_0 + \sum_{k=1}^{K}x_k^T\beta_k \\
\beta_0, \beta &\sim \text{Normal}(0, 1)
\end{aligned}
\]

where \(\beta_0\) is the intercept and \(E_i\) is an offset term.

The offset \(E_i\) is often termed ``exposure'' in applications such as
disease risk modeling where the offset may be the number of observed
cases in a previous year, for example. More technically, the offset term
scales the Poisson output to be a rate rather than a count. This is
appropriate when observations \(i\) have different potentials for
response \(Y\). For example, a county with higher population will
naturally have a higher count of patients with asthma than a county with
a smaller population. Mathematically, the offset is a covariate with
parameter set equal to 1. Recall, the Poisson regression assumes we can
model the mean of response \(Y\), \(\bar{Y}\) with a combination of
linear predictors:

\[
log(\bar{Y}) = \beta_0 + \sum_{k=1}^{K}x_k^T\beta_k \\
\] Therefore, if we wish to model the rate \(Y/E\), the equation is
rewritten

\[
\begin{aligned}
log(\bar{Y}/E) & = \beta^{'}_0 + \sum_{k=1}^{K}x_k^T\beta^{'}_k \\
log(\bar{Y}) & = log(E) + \beta^{'}_0 + \sum_{k=1}^{K}x_k^T\beta^{'}_k
\end{aligned}
\]

In this study, we define \(E_i\) to equal the number of times a bus
stops in block group \(i\). This allows us to control for block groups
with more or less supply.

In the basic Poisson regression, we often give \(\beta_0\) and \(\beta\)
vague priors such as Normal(\(0, 1\)). The exposure term \(E_i\) comes
from the data and does not recieve a prior.

\subsubsection{Model 1}\label{model-1}

We will call this baseline Poisson regression Model 1. Model 1 uses all
18 covariates with simple Normal(0, 1) priors and no overdispersion
parameters. The Stan program used to fit Model 1, as well as the other
four models, can be found in the appendix. Model 1 fits quickly and
efficiently. The Stan output for \(\beta_0\) and \(\beta\) is below.

@@

Model 1 produces coefficient estimates with tight confidence intervals;
few of the \(\beta_k\) 95\% intervals cross zero. Based on initial
diagnostics and precise parameter estimates, Model 1 appears to be a
``good'' fit. In this way, Model 1 illustrates the importance of
diagnostics beyond the default Stan output.

@@

\subsection{Overdispersed Poisson
Regression}\label{overdispersed-poisson-regression}

The most obvious assumption made in a Poisson regression is that
response variable \(Y\) follows a Poisson distribution. In addition to
requiring integer counts, this assumption requires that
\(\text{Var}(Y) = \text{E}(Y)\). This is often not true in practice. In
the case of the Metro Transit ridership data,
\(\text{Var}(Y) >>\text{E}(Y)\). In this case, we call the unexpectedly
large variance ``overdispersion'' or ``extra-Poisson variance.''

@@

\[
\begin{aligned}
\eta_i = log(\lambda_i) & = \beta_0 + \sum_{k=1}^{K}x_k^T\beta_k + \theta\sigma \\
\theta &\sim \text{Normal}(0, 1) \\
\sigma &\sim \text{Normal}(0, 5)
\end{aligned}
\]

where \(\theta\) is a set of heterogenous random effects. In practice,
\(\theta\) is scaled by its variance parameter \(\sigma\) to allow
eassier fitting in Stan. The addition of this set of random effects
improves fits on data with overdispersion. With sufficiently vague
priors, \(\theta\) and \(\sigma\) can account for most or all of the
overdispersion not accounted for by
\(\beta_0 + \sum_{k=1}^{K}x_k^T\beta_k\).

\subsubsection{Model 2}\label{model-2}

Model 2 builds upon Model 1 by incorporating the set of random effects
described above. Model 2 is \emph{significantly} more computationally
expensive to fit than Model 1. This is due to relatively large
autocorrelation between each iteration, visible in the traceplots below
(Figure \_\_\_). This autocorrelation made it necessary to run Model 2
on four chains for 20,000 iterations each. Despite this, \texttt{n\_eff}
is significantly lower for Model 2 than Model 1. Model 2 also
experienced 2 transitions exceeding maixmum treedepth: not a problem in
terms of model fit, but a signal that the posterior is more challenging
to fit. Reparameterization could likely reduce autocorrelation between
iterations.

The parameter estimates for \(\beta_0\) and \(\beta\) for Model 2 are
below. Notably, the direction and magnitude of the coefficient estimates
are similar to those produced by Model 1. However, the coverage
intervals are much wider. Based on ``test runs'' of Model 2, we expect
increasing the number of iterations to decrease the width of these
coverage intervals (we observe that the converage interval is
proportional to the number of iterations in this setting).

@@

Because we gave \(\theta\) and \(\sigma\) vague priors, we can see a
\emph{much} better fit in the posterior predictive checks for Model 2.
One \emph{could} stop analysis here. However, the addition of
\(\theta \sigma\) only improves the \emph{fit} of the model, not our
understanding of the data. Therefore, for this study, we will decompose
\(\theta \sigma\) into spatial and non-spatial random effects. First,
however, we will implement a Bayesian variable selection process to
determine which, if any, of our covariates \(x_k\) should be removed
from the analysis.

\subsection{Horseshoe Priors and Variable
Selection}\label{horseshoe-priors-and-variable-selection}

In the baseline model, we give coefficients \(\beta_1, ..., \beta_k\)
simple normal priors. With a large set of possible covariates, however,
we may reasonably expect some of the \(\beta_k\) coefficients to be
equal to zero -- in other words, that some of the parameters \(x_k\)
have no effect on response \(Y\). In frequentist applications, we might
use a method such as the LASSO (Least Absolute Shrinkage and Selection
Operator) to identify relevant variables. There are several Bayesian
alternatives for variable selection with generally consist of applying
particular priors to \(\beta\).

\subsubsection{Spike and Slab - needs
graphics}\label{spike-and-slab---needs-graphics}

In the Bayesian setting, there are two primary methods for variable
selection. The first is called the ``spike-and-slab'' prior (Mitchell
and Beauchamp, 1988; George and McCulloch, 1999) and has often been
considered the ``gold standard'' for sparse Bayesian regression, or
variable selection (Piironen and Vehtari, 2017). This prior is often
expressed as

\[
\begin{aligned}
\beta_j | \lambda_j, c, \epsilon & \sim \lambda_j \text{Normal}(0, c^2) + (1-\lambda_j)\text{Normal}(0, \epsilon^2) \\
\lambda_j & \sim \text{Bernoulli}(\pi)
\end{aligned}
\]

for \(j = 1, 2, ..., J\) where \(\lambda_j \in {0, 1}\) indicates
whether \(\beta_j\) is from the ``spike'' (ie, near zero) or the
``slab'' (ie, nonzero). The ``spike'' is the area where most of the
prior density for \(\beta\) is centered (generally around 0) while the
``slab'' refers to the low-density extent of the prior. The
spike-and-slab prior encourages all \(\beta_j\) towards zero; only
\(\beta_j\) sufficiently far from 0 will be estimated to be from the
slab.

The primary challenge with the spike-and-slab (and other Bayesian
shrinkage methods) is the selction of priors. Priors must be set for the
width of the slab \(c\) and the prior inclusion probability \(\pi\).
Here, \(\pi\) reflects our prior understanding of the sparsity of
\(\beta\). In practice, we rarely have strong prior knowledge of the
number of predictors we expect to be distinguishable from zero. This can
make the implementation of sparse regression more challenging than, say,
the frequentist LASSO.

\subsubsection{Horseshoe - needs
graphics}\label{horseshoe---needs-graphics}

The second method for Bayesian sparse regression is to give \(\beta\)
some continuous \emph{sparsity inducing prior}. \textbf{Van Erp et al
(2019)} provide an introduction to many such continous priors. Ridge
regression and the Bayesian LASSO are two of the most approachable
priors, but all sparsity inducing priors utilize the same logic as the
spike-and-slab: give \(\beta\) a prior with \emph{most} of its mass near
0, to shrink irrelevant coefficients, and the rest of its mass ``far''
from 0. Identifying ``far'' depends more or less on the specific data
and model in question.

The horseshoe prior, proposed by \textbf{Carvalho et al, 2010} is a
popular choice in Bayesian literature, in part because it is similar to
the ``gold standard'' spike-and-slab. Following the notation of
\textbf{Vehtari \& Piironen}, the horseshoe prior can be expressed

\[
\begin{aligned}
\beta_k \text{ | } \lambda_k, \tau & \sim \text{Normal}(0, \tau^2\lambda_k^2) \\
\lambda_k & \sim \text{Half-Cauchy}(0, 1)
\end{aligned}
\]

for \(k = 1, 2, ..., K\). The horseshoe prior is so-named because for
fixed values \(\tau = \lambda_k = 1\), the prior resemebles a Beta(1/2,
1/2) distribution, or a horseshoe.

The horseshoe is often favored because it is a global-local shrinkage
prior. This means that \(\tau\) shrinks \emph{all} parameters towards 0
while the local parameter \(\lambda_k\) and the heavy Cauchy tails allow
larger coefficients to remain unshrunk. While this is precisely the goal
of a sparsity inducing prior, the horseshoe prior can fail to regularize
large coefficients \emph{at all}, which can be a problem when parameters
are weakly identified by data. With no regularization applied to the
largest coefficients, there is a risk of overfitting. Additionally,
there is no consensus in the literature for assigning priors to \(tau\).
Piironen and Vehtari (2017) introduced the \emph{regularized horseshoe}
to address both of these shortcomings of the horseshoe.

\subsubsection{Regularized Horseshoe - needs
graphics}\label{regularized-horseshoe---needs-graphics}
